### 2021-06-08 paper/report.tex
Updated first few paragraphs.

Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity
on many dimensions. The development of highly effective vaccines holds the promise of
containment in the medium term. However, most countries find themselves many
months---and often years---away from reaching vaccination-induced herd
immunity.\comment[id=HM]{Cite some paper on herd immunity, maybe vaccine data} In the
meantime, it is of utmost importance to employ an effective mix of strategies for
containing the virus. The most frequent initial response was a set of non-pharmaceutical
interventions (NPIs) to reduce contacts between individuals. While this has allowed some
countries to sustain equilibria with very low infection numbers\footnote{See
\citet{Contreras2021} for a theoretical equilibrium at low case numbers which is
sustained with test-trace-and-isolate policies.}, most have seen large ups and downs in
their infection rates. Containment measures have become increasingly diverse and
included testing, more nuanced NPIs, and contact tracing. Neither these policies' effect
nor the influence of seasonal patterns or more infectious virus strains are well
understood in quantitative terms. This paper develops a model incorporating all these
factors. The framework allows to combine a wide variety of data in a timely fashion,
making it useful to predict the effects of various interventions. We apply the model to
Germany and show that rapid testing had the largest impact on the reduction in
infections by XXX\% during the A weeks between X April and Y May.\comment[id=HM]{insert
some useful numbers}. We conclude that rapid tests should play a large role...

At the core of our agent-based model are physical contacts between heterogeneous agents
(Figure~\ref{fig:broad_model_description}).\footnote{We provide a detailed comparison to
other approaches in \ref{sec:literature_review}. The model most closely related to ours
is described in \citet{Hinch2020}.} Each contact between an infectious individual and
somebody susceptible to the disease bears the risk of transmitting the virus. Contacts
occur in the household, at work, at school, or in other settings (leisure activities,
grocery shopping, medical appointments, etc.).\comment[id=HM]{Systemic relevance of
work?} Some contacts recur regularly, others occur at random. Random contacts are
typically assortative in age and geographical location. Empirical applications can take
the population structure from census data and the types and frequency of contacts from
diary data measuring contacts before the pandemic
\citep[e.g.][]{Mossong2008}.\footnote{\citet{Hoang2019} provide access to multiple data
sets on contact types and frequencies at \url{http://www.socialcontactdata.org/}
covering countries from all continents except North America and Australia.} The
dimensions are chosen so that the most common NPIs can be modeled in great detail by
reducing the number of contacts in a particular setting or the risk of transmitting the
disease for a type of contact. For example, a mandate to work from home will reduce the
number of work contacts to zero for a fraction of the working population. Schools and
daycare can be closed entirely, operate at reduced capacity including an alternating
schedule, or implement mitigation measures like masking requirements or air filters
\citep{Lessler2021}. Curfews may reduce the number of contacts in non-work/non-school
settings. In any setting, measures like masking requirements would reduce the
probability of infection associated with a contact.

@@

    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/model-graph-top-left}
        \caption{{\small Model description}}
        \label{fig:broad_model_description}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/model-graph-top-right}
        \caption{Disease progression}
        \label{fig:disease_progression}
    \end{subfigure}
    \vskip3ex
    \begin{subfigure}[b]{0.425\textwidth}
        \centering

        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-left}
        \caption{{\small PCR and antigen tests}}
        \label{fig:pcr_antigen_tests}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering

        Model for detected/undetected cases

        % \includegraphics[width=\textwidth]{../figures/model_detailed.png}
        \caption{{\small Model for translating simulated number to officially recorded
        cases}}
        \label{fig:model_for_official_cases}
    \end{subfigure}

@@

The model includes several other features, which are crucial to describe the evolution of
the pandemic in 2020-2021. New virus strains with different profiles regarding
infectiousness \replaced[id=K]{}{and disease progress}\comment[id=K]{sid only supports
differing infectiousness between variants at the moment} can be introduced. With a
probability of 75\% \citep{Hunter2021}, vaccinated agents become immune and they do not
transmit the virus \citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.\footnote{75\%
is lower than what is usually reported for after the second dose. We choose it because in
our model does not include booster shots and does not allow for vaccinated individuals to
become infected and transmit the disease which is the case for Covid-19
\citep{Petter2021, LevineTiefenbrun2021, Pritchard2021}.}% the sources only report reduced viral shedding and reduced risk of infection.
% Lauterbach calculated that taking the risk of infection * reduced infectiousness given
% infection based on theses studies suggest that given exposure the risk that fully
% vaccinated individuals pose is <10% of an unvaccinated person.
While vaccines are rolled out, priority depends on age and occupation. Agents may
demand two types of tests as described in Figure~\ref{fig:pcr_antigen_tests}. First,
polymerase chain reaction (PCR) tests directly reveal whether an individual is infected
or not. PCR tests require some time to be processed and there are always aggregate
capacity constraints. Second, rapid antigen tests yield immediate results. Specificity
and sensitivity of these tests is set according to data by \cite{Bruemmer2021,
Smith2021} . The sensitivity of the rapid antigen tests depends on the timing of the
test relative to the start of infectiousness.\comment[id=HM]{subsequent PCR tests,
quarantine, work/school}. Demand for rapid tests will depend on policy, e.g., prices,
testing requirements before work, attending school, or visiting restaurants and shops.

>>>>

Since early 2020, the CoViD-19 pandemic has presented an enormous challenge to humanity
on many dimensions. The development of highly effective vaccines holds the promise of
containment in the medium term. However, most countries find themselves many
months---and often years---away from reaching vaccination-induced herd
immunity.\comment[id=HM]{Cite some paper on herd immunity, maybe vaccine data} In the
meantime, it is of utmost importance to employ an effective mix of strategies for
containing the virus. The most frequent initial response was a set of non-pharmaceutical
interventions (NPIs) to reduce contacts between individuals. While this has allowed some
countries to sustain equilibria with very low infection numbers\footnote{See
\citet{Contreras2021} for a theoretical equilibrium at low case numbers which is
sustained with test-trace-and-isolate policies.}, most have seen large ups and downs in
their infection rates. Containment measures have become increasingly diverse and
included testing, more nuanced NPIs, and contact tracing. Neither these policies' effect
nor the influence of seasonal patterns or more infectious virus strains are well
understood in quantitative terms. This paper develops a model incorporating all these
factors. The framework allows to combine a wide variety of data in a timely fashion,
making it useful to predict the effects of various interventions. We apply the model to
Germany and show that rapid testing had the largest impact on the reduction in
infections by almost 80\% during the month of May 2021. We conclude that rapid tests
have a large role to play at least as long as vaccinations have not been offered to an
entire population.

At the core of our agent-based model are physical contacts between heterogeneous agents
(Figure~\ref{fig:broad_model_description}).\footnote{We provide a detailed comparison to
other approaches in \ref{sec:literature_review}. The model most closely related to ours
is described in \citet{Hinch2020}.} Each contact between an infectious individual and
somebody susceptible to the disease bears the risk of transmitting the virus. Contacts
occur in the household, at work, at school, or in other settings (leisure activities,
grocery shopping, medical appointments, etc.). Some contacts recur regularly, others
occur at random. Random contacts are typically assortative in age and geographical
location. Empirical applications can take the population structure from census data and
the types and frequency of contacts from diary data measuring contacts before the
pandemic \citep[e.g.][]{Mossong2008}.\footnote{\citet{Hoang2019} provide access to
multiple data sets on contact types and frequencies at
\url{http://www.socialcontactdata.org/} covering countries from all continents except
North America and Australia.} The dimensions are chosen so that the most common NPIs can
be modeled in great detail by reducing the number of contacts in a particular setting or
the risk of transmitting the disease for a type of contact. For example, a mandate to
work from home will reduce the number of work contacts to zero for a fraction of the
working population. Schools and daycare can be closed entirely, operate at reduced
capacity---including an alternating schedule---,or implement mitigation measures like
masking requirements or air filters \citep{Lessler2021}. Curfews may reduce the number
of contacts in non-work/non-school settings. In any setting, measures like masking
requirements would reduce the probability of infection associated with a contact.

@@

    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/model-graph-top-left}
        \caption{{\small Model description}}
        \label{fig:broad_model_description}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../figures/model-graph-top-right}
        \vskip4ex

        \caption{Disease progression}
        \label{fig:disease_progression}
    \end{subfigure}
    \vskip3ex
    \begin{subfigure}[b]{0.425\textwidth}
        \centering

        \includegraphics[width=\textwidth]{../figures/model-graph-bottom-left}
        \caption{{\small Model for antigen tests}}
        \label{fig:antigen_tests}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.425\textwidth}
        \centering
        % \includegraphics[width=\textwidth]{../figures/model_detailed.png}
        Model for undetected cases
        \vskip4ex

        \caption{{\small Model for undetected cases}}
        \label{fig:model_for_official_cases}
    \end{subfigure}

@@

The model includes several other features, which are crucial to describe the evolution
of the pandemic in 2020-2021. New virus strains with different profiles regarding
infectiousness \replaced[id=K]{}{and disease progress}\comment[id=K]{sid only supports
differing infectiousness between variants at the moment} can be introduced. With a
probability of 75\% \citep{Hunter2021}, vaccinated agents become immune and they do not
transmit the virus \citep{Petter2021, LevineTiefenbrun2021,
Pritchard2021}.\footnote{75\% is lower than what is usually reported for after the
second dose. We choose it because in our model does not include booster shots and does
not allow for vaccinated individuals to become infected and transmit the disease which
is the case for Covid-19 \citep{Petter2021, LevineTiefenbrun2021,
Pritchard2021}.}% the sources only report reduced viral shedding and reduced risk of infection.
% Lauterbach calculated that taking the risk of infection * reduced infectiousness given
% infection based on theses studies suggest that given exposure the risk that fully
% vaccinated individuals pose is <10% of an unvaccinated person.
While vaccines are rolled out, priority depends on age and occupation. Agents may demand
two types of tests as described in Figure~\ref{fig:antigen_tests}. First, polymerase
chain reaction (PCR) tests directly reveal whether an individual is infected or not. PCR
tests require some time to be processed and there are always aggregate capacity
constraints. Second, rapid antigen tests yield immediate results. Specificity and
sensitivity of these tests is set according to data by \cite{Bruemmer2021, Smith2021} .
The sensitivity of the rapid antigen tests depends on the timing of the test relative to
the start of infectiousness.\comment[id=HM]{subsequent PCR tests, quarantine,
work/school}. Demand for rapid tests will depend on policy, e.g., prices, testing
requirements before work, attending school, or visiting restaurants and shops.
